---
title: "Code for Task 1.1 - data collection - BLIZ"
author: "Leticia Duarte"
output: 
  html_notebook:
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

<style type="text/css">

body{ /* Normal  */
      font-size: 13px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 20px;
  font-weight: bold;
}
h1 { /* Header 1 */
  font-size: 20px;
  font-weight: bold;
}
h2 { /* Header 2 */
    font-size: 17px;
  font-weight: bold;
}
h3 { /* Header 3 */
    font-size: 15px;
  font-weight: bold;
}
code.r{ /* Code block */
    font-size: 12px;
}
pre{ /* Code block - determines code spacing between lines */
    font-size: 12px;
}
h1, .h1, h2, .h2, h3, .h3, h4, .h4 {
    margin-top: 24xpx;
}
</style>


# Summary of the analyses

In **Step 1**, we prepared the candidate species for both Bavaria and Baden-Württemberg (BW).

- For Bavaria we used the species found in https://www.lfu.bayern.de/natur/rote_liste_pflanzen/doc/pflanzen/rl_pflanzen_nach_kategorien.pdf

- For both Bavaria and BW we went to https://www.gbif.org/occurrence/search?occurrence_status=present&q= to download the species list with occurrences, and used the following filters: 
  - "Scientific name" - "Tracheophyta"
  - "Location" - "Including coordinates"
  - "Administrative areas" - "Bayern" or Baden-Württemberg"
  

In **Step 2** we looked for the available traits and checked which species had information for all traits.

For each trait we did the following steps:

- Download data
- Check units and methods used, and filter the useful data
- Select the columns with the measurements, methods and units
- Format data
- Merge data with the candidate species list prepared in Step 1

After checking and preparing all traits (Seed number, Germination rate, Carrying capacity, Wetmass, and Dispersal distance), we joined them, in order to check which species had information for them most of them. Because carrying capacity was a limiting variable, we decided to use the mean of the data from species of the same genus for the species that had missing this data. We did the same for the Germination rate for 2 species. We calculated the Mortality rate using an allometric equation based on the values of wetmass.

Finally, we formatted all the data to compile information for the chosen species in one table, and calculated the mean and standard deviation for each.


In **Step 3**, we calculated the number of occurrences for the chosen species.


In **Step 4**, we mapped the occurrence data of the chosen species


Detailed information about each step can be found in the code below.

<span style="color: darkorange;">All data files necessary can be found in https://drive.google.com/drive/folders/1x93UjAnduPXwROqXVnQeJ4JQbdVoKeom?usp=sharing


# 1 - Candidate species


## Species from Bavaria

Download red list of vascular plants from Bavaria (https://www.lfu.bayern.de/natur/rote_liste_pflanzen/doc/pflanzen/rl_pflanzen_nach_kategorien.pdf) and convert to csv

```{r}

plants_bavaria <- readLines("Data/Rote_List_Bayern.csv") #read csv as a text object

simplified_plants = paste(plants_bavaria, collapse = " ") #simplify data by collapsing everything with a space as delimiter

#run a for loop where we are looking for every "space + capital letter" and want to replace with a "new line + capital letter" - this was necessary because we had multiple species listed in one row
for (x in LETTERS) {
simplified_plants = gsub(paste0(" ",x), paste0("\n",x), simplified_plants)
}

#formatting data
sp_bavaria_red = 
  simplified_plants |>
  strsplit(split = "\n") |> #split lines in substrings
  unlist() |>
  gsub(paste0(" ","´"), paste0(""), x = _) |> #fix text (eliminate "´")
  gsub(paste0(" ","x"), paste0(""), x = _) |> #fix text (eliminate "x")
  unique() #remove duplicates

sp_bavaria_red <- data.frame(sp_bavaria_red) #transform into dataframe
colnames(sp_bavaria_red) <- c('species') #change name of id column

```


Download list of species with occurrences in Bavaria from GBIF database (https://www.gbif.org/occurrence/search?occurrence_status=present&q=), and use the following filters: 
  - "Scientific name" - "Tracheophyta"
  - "Location" - "Including coordinates"
  - "Administrative areas" - "Bayern"
  
Download - option "species list"
  
(All occurrence data in BIEN is from GBIF, so there was no need to use both databases)
  
```{r}

GBIF_bav <- read.csv("Data/SpeciesList_Bavaria.csv", sep = "\t") #read file

sp_bav <- GBIF_bav[c("species")] #select column with species names
sp_bav <- sp_bav[sp_bav$species != "", , drop=FALSE] #eliminate empty rows

```


## Species from Baden-Württemberg (BW)

Download list of species with occurrences in BW from GBIF database (https://www.gbif.org/occurrence/search?occurrence_status=present&q=), and use the following filters: 
  - "Scientific name" - "Tracheophyta"
  - "Location" - "Including coordinates"
  - "Administrative areas" - "Baden-Württemberg"
  
Download - option "species list"

```{r}

GBIF_baden <- read.csv("Data/SpeciesList_Baden.csv", sep = "\t") #read file

sp_baden <- GBIF_baden[c("species")] #select column with species names
sp_baden <- sp_baden[sp_baden$species != "", , drop=FALSE]#eliminate empty rows

```

## Candidate species list

```{r}

GBIF_Species <- merge(sp_bav,sp_baden, by = "species", all = TRUE) #merge both lists from GBIF
BB_Species <- merge(GBIF_Species,sp_bavaria_red, by = "species", all = TRUE) #merge GBIF data with Red List data

length(unique(BB_Species$species)) #calculate number of candidate species

```

# 2 - Traits

## Number of Seeds - reproductive rate

Download "Seed number" file from LEDA database (https://uol.de/en/landeco/research/leda/data-files)
Files (.txt) from this database have formatting problems - need to remove manually the first few rows to be read

**Brief summary:**
Filtered only values with **reproduction.unit.measured = 'per ramet/tussock or individual plant'. For the methods used, considered all of them

```{r}

seed_number <- read.csv2("Data/seed_number.txt", quote = "", dec = ".") #read file

unique(seed_number$reproduction.unit.measured) #check units measured
unique(seed_number$general.method) #check methods used

SeedNumber <- subset(seed_number, reproduction.unit.measured == 'per ramet/tussock or individual plant ') #filter rows with reproduction.unit.measured = 'per ramet/tussock or individual plant '
SeedNumber <- SeedNumber[c("SBS.name","single.value","general.method","reproduction.unit.measured")] #select columns with species, values, methods and units
names(SeedNumber)[names(SeedNumber) == 'SBS.name'] <- 'species' #change column name
names(SeedNumber)[names(SeedNumber) == 'single.value'] <- 'seed_number' #change column name

str(SeedNumber) #check if values are all numeric

merge_seeds <- merge(BB_Species, SeedNumber, by = "species") #merge data with candidate species list

length(unique(merge_seeds$species)) #check the number of species with data for this trait

```
## Germination rate

Germination rate is calculated by dividing the number of seeds that sprout by the total number of seeds, and then multiplying that number by 100.

Download "95 - Seed germination rate" from TRY database (https://www.try-db.org/TryWeb/Home.php)

**Brief summary:**
Most if not all data for this trait is based in controlled laboratory experiments &rarr; <span style="color: red;"> can we use this data? </span>
Germination rate in ideal conditions.


```{r}

germination_rate <- read.delim("Data/SeedGermination.txt") #read file

SeedGermination <- subset(germination_rate, TraitID == 95) # filter rows with germination rate values
SeedGermination <- SeedGermination[SeedGermination$StdValue != "", , drop=FALSE] #eliminate empty rows

unique(SeedGermination$UnitName) #check units measured
unique(SeedGermination$OriglName) #check methods used

SeedGermination <- SeedGermination[c("AccSpeciesName","OriglName","StdValue","UnitName")] #select columns with species, values, methods and units
names(SeedGermination)[names(SeedGermination) == 'AccSpeciesName'] <- 'species' #change name
names(SeedGermination)[names(SeedGermination) == 'OriglName'] <- 'germination_method' #change name
names(SeedGermination)[names(SeedGermination) == 'StdValue'] <- 'germination_perc' #change name
names(SeedGermination)[names(SeedGermination) == 'UnitName'] <- 'germination_unit' #change name

str(SeedGermination) #check if values are all numeric

merge_germination_rate <- merge(BB_Species, SeedGermination, by = "species") #merge data with candidate species list

```

### Add genus data for species missing

Because there are some data missing for some species, we will use the values from other species of the same genus.

```{r}
genus_list_ger <- read.csv("Data/genus_list_Germination.csv") #read file

mean_germination_sp <- aggregate(germination_perc~species, data=SeedGermination, mean) #calculate the mean germination rate for each species, because there are multiple values

for (row in seq_len(nrow(genus_list_ger))) { #for each row in the genus list
  curr_genus = genus_list_ger[row, "genus"] #get each genus
  curr_sp = genus_list_ger[row, "sp"] #get each species
  
  curr_genus_subset = mean_germination_sp[grep(curr_genus, mean_germination_sp$species), ] #create a subset of data for the current genus
  curr_genus_nrows = nrow(curr_genus_subset)
  
  curr_df = data.frame(species = rep(curr_sp, curr_genus_nrows), germination_method = rep("%Germination", curr_genus_nrows), germinaton_perc = curr_genus_subset$germination_perc, germination_unit = rep("%", curr_genus_nrows)) #create dataframe with new data
  names(curr_df) <- names(SeedGermination) #make sure column names are the same as the original dataframe
  
  SeedGermination <- rbind(SeedGermination, curr_df) #merge new data with original dataframe
}


```


## Carrying capacity
= maximum planting density

Download "601 - Species tolerance to planting density" from TRY database (https://www.try-db.org/TryWeb/Home.php)

This same data exists in FloraWeb.

reference: The PLANTS Database, United States Department of Agriculture, National Resources Conservation Service. http://plants.usda.gov/
measurement remarks:
These data have been gathered from the scientific literature, gray literature, agency documents, and the knowledge of plant specialists. Characteristics data values are best viewed as approximations since they are primarily based on field observations and estimates from the literature, not precise measurements or experiments. Characteristics for the many conservation plant species native to the U.S. were typically provided by experts familiar with the species in its natural setting. Most values given apply to plants nationwide. Many values are relative to other species since absolute figures are not available.


```{r}

density <- read.delim("Data/planting_density.txt") #read file

PDensity <- subset(density, DataID == 1320) # filter rows with maximum planting density - the database has other information

unique(PDensity$OrigUnitStr) #check units measured

PDensity <- PDensity[c("AccSpeciesName","OriglName","OrigValueStr","OrigUnitStr")] #select columns with species, values, methods and units
names(PDensity)[names(PDensity) == 'AccSpeciesName'] <- 'species' #change name
names(PDensity)[names(PDensity) == 'OriglName'] <- 'density_method' #change name
names(PDensity)[names(PDensity) == 'OrigValueStr'] <- 'density_acre' #change name
names(PDensity)[names(PDensity) == 'OrigUnitStr'] <- 'density_unit' #change name

str(PDensity) #check if values are all numeric

merge_density <- merge(BB_Species, PDensity, by = "species") #merge data with candidate species list

```

### Add genus data for species with missing data

Because there are some data missing for some species, we will use the values from other species of the same genus.

```{r}
genus_list_den <- read.csv2("Data/genus_list_PDensity.csv") #read file

for (row in seq_len(nrow(genus_list_den))) { #for each row in the genus list
  curr_genus = genus_list_den[row, "genus"] #get each genus
  curr_sp = genus_list_den[row, "sp"] #get each species
  
  curr_genus_subset = PDensity[grep(curr_genus, PDensity$species), ] #create a subset of data for the current genus
  curr_genus_nrows = nrow(curr_genus_subset)
  
  curr_df = data.frame(species = rep(curr_sp, curr_genus_nrows), density_method = rep("Planting Density per Acre, Maximum", curr_genus_nrows), density_acre = curr_genus_subset$density_acre, density_unit = rep("1/acre", curr_genus_nrows)) #create dataframe with new data
  names(curr_df) <- names(PDensity) #make sure column names are the same as the original dataframe
  
  PDensity <- rbind(PDensity, curr_df) #merge new data with original dataframe
}


#convert plants/acre to plants/m2
# 1 acre = 4046.86 m2
PDensity$density_sqMeter = PDensity$density_acre/4046.86

```


## Wetmass

Download "3025 - plant wet biomass" from TRY database (https://www.try-db.org/TryWeb/Home.php)

Brief summary:
Didn't need to filter units or methods because they were all the same (g and whole plant fresh mass).

```{r}

plant_fresh_mass <- read.delim("Data/PlantFreshMass.txt") #read file

PlantFreshMass <- subset(plant_fresh_mass, TraitID == 3407) #filter rows with fresh mass values - the database has other information

unique(PlantFreshMass$OrigUnitStr) #check units measured
unique(PlantFreshMass$DataName) #check methods used

PlantFreshMass <- PlantFreshMass[c("AccSpeciesName","DataName","OrigValueStr","OrigUnitStr")] #select columns with species, values, methods and units
names(PlantFreshMass)[names(PlantFreshMass) == 'AccSpeciesName'] <- 'species' #change name
names(PlantFreshMass)[names(PlantFreshMass) == 'DataName'] <- 'P_part' #change name
names(PlantFreshMass)[names(PlantFreshMass) == 'OrigValueStr'] <- 'wetmass_g' #change name
names(PlantFreshMass)[names(PlantFreshMass) == 'OrigUnitStr'] <- 'Unit' #change name

str(PlantFreshMass) #check if values are all numeric
#if not numeric, go check if any values are wrong (no problems)
PlantFreshMass$wetmass_g <- as.numeric(PlantFreshMass$wetmass_g) #convert column with values to numeric instead of character

merge_fresh_mass <- merge(BB_Species, PlantFreshMass, by = "species") #merge data with candidate species list

```


Download data from Supporting Information of "PlantNE: a global database of plant biomass from nitrogen-addition experiments" (https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.2840)

**Brief summary:**
Used the biomass data from the control groups used in this study (column CK_M).
Filtered data with unit "g" and "g " because they are the same, and to be consistent.

```{r}

plantNE_data <- read.csv("Data/PlantNE_Biomass.csv", dec = ".") #read file

unique(plantNE_data$Unit) #check units measured
unique(plantNE_data$P_part) #check part of the plant measured

plantNE <- subset(plantNE_data, Unit %in% c('g','g ') & P_part == 'WP') #filter rows with unit = "g" and plant part = "whole plant"
plantNE <- plantNE[c("Species","CK_M","Unit","P_part")] #select columns with species, values, methods and units
names(plantNE)[names(plantNE) == 'Species'] <- 'species' #change name
names(plantNE)[names(plantNE) == 'CK_M'] <- 'wetmass_g' #change name

str(plantNE) #check if values are all numeric

merge_plantNE <- merge(BB_Species, plantNE, by = "species") #merge data with candidate species list

```

Join data from both databases

```{r}

Wetmass <- rbind(PlantFreshMass, plantNE)

length(unique(Wetmass$species)) #check the number of species with data for this trait

```

## Mortality rate

Mortality rate was calculated from biomass with the allometric equation from Marba, N., Duarte, C.M. & Agustí S. (2007) Allometric scaling of plant life history. Proceedings of the National Academy of Sciences of the United States of America, 104, 15777–15780. (https://www.pnas.org/doi/abs/10.1073/pnas.0703476104)

“Marba et al. (2007)  applied allometric theory to the relationship between size and important life-history parameters and reported that mortality and birth rates scaled as **the –0.25 power and life span as the 0.25 power of plant mass from phytoplankton to trees**.” (https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/j.1365-2745.2008.01415.x) 


Equation: wetmass^-0.25


```{r}

Wetmass$mortality <- Wetmass$wetmass_g^-0.25 #mortality equation by Marba et al. (2007)

Wetmass <- subset(Wetmass, !wetmass_g<1) #exclude species with wetmass < 1g

```


## Dispersal Distance

Download data from the Supporting Information of "Seed dispersal distance classes and dispersal modes for the European flora" (https://onlinelibrary.wiley.com/doi/full/10.1111/geb.13712)

**This article has the data in dispersal classes:**

Dispersal distance classes from Vittoz, P., & Engler, R. (2007). Seed dispersal distances: A typology based on dispersal modes and plant traits. Botanica Helvetica, 117, 109–124.

(dispersal distances in m for 50 and 99% seeds)

- class 1: 0.1-1 m

- class 2: 1-5 m

- class 3: 2-15 m

- class 4: 40-150 m

- class 5: 10-500 m

- class 6: 400-1500 m 

- (class 7: 500-5000 m - **excluded this class because it's dispersal by humans and that data exists only for some species**)

**&rarr; Used the 99% distribution kernels as the maximum distance**

**&rarr; Used the 50% distribution kernels as the mean distance**

```{r}

dispersal <- read.csv2("Data/Lososova_et_al_2023_Dispersal.csv") #read file

Dispersal <- dispersal[c("Taxon","Dispersal.distance.class..1.6.")] #select columns with species and dispersal class
names(Dispersal)[names(Dispersal) == 'Taxon'] <- 'species' #change name
names(Dispersal)[names(Dispersal) == 'Dispersal.distance.class..1.6.'] <- 'dispersal_class_1-6' #change name

##convert class data in mean distance and max distance
dispersal_key = read.csv("Data/dispersal_classes.csv") #read file with dispersal classes and respective mean and max values

#assign max and mean values for each row
for (row in seq_len(nrow(dispersal_key))) {
    curr_class = dispersal_key[row, "class"]
    curr_mean = dispersal_key[row, "mean_dist_m"]
    curr_max = dispersal_key[row, "max_dist_m"]
    
    Dispersal$mean_dist_m[Dispersal$'dispersal_class_1-6' == curr_class] = curr_mean
    Dispersal$max_dist_m[Dispersal$'dispersal_class_1-6' == curr_class] = curr_max
}

str(Dispersal) #check if values are all numeric

merge_dispersal <- merge(BB_Species, Dispersal, by = "species") #merge data with candidate species list

length(unique(merge_dispersal$species)) #check the number of species with data for this trait


```

## Lambda

```{r}
library(Rcompadre)

load("Data/COMPADRE_v.6.23.5.0.RData") #load compadre database
compadre <- as_cdb(compadre) #ensure that the database is of the correct class

genus_list_Lambda <- read.csv("Data/genus_list_Lambda.csv") #list with species and genus to look for

sub_genus <- subset(compadre, Genus %in% genus_list_Lambda$genus) #subset database for our genus
unique(sub_genus$MatrixTreatment) #check methods

sub_genus2 <- subset(sub_genus, MatrixTreatment == "Unmanipulated") #filter rows with only unmanipulated values
unique(sub_genus2$Genus) #check genus with information
unique(sub_genus2$SpeciesAccepted) #check species with information

lambdaVals <- sapply(matA(sub_genus2), popdemo::eigs, what = "lambda") #calculate lambda values

Lambda <- data.frame(sub_genus2$SpeciesAccepted, lambdaVals) #create dataframe with lambda values

names(Lambda)[names(Lambda) == 'sub_genus2.SpeciesAccepted'] <- 'species' #change name

```

### Add genus data for species missing

Because there are some data missing for some species, we will use the values from other species of the same genus.

```{r}

for (row in seq_len(nrow(genus_list_Lambda))) { #for each row in the genus list
  curr_genus = genus_list_Lambda[row, "genus"] #get each genus
  curr_sp = genus_list_Lambda[row, "sp"] #get each species
  
  curr_genus_subset = Lambda[grep(curr_genus, Lambda$species), ] #create a subset of data for the current genus
  curr_genus_nrows = nrow(curr_genus_subset)
  
  curr_df = data.frame(species = rep(curr_sp, curr_genus_nrows), lambdaVals = curr_genus_subset$lambdaVals) #create dataframe with new data
  
  Lambda <- rbind(Lambda, curr_df) #merge new data with original dataframe
}

```


# 2.1 - Merge all traits

Merge all trait data and check which species have data for all of them.

```{r}

traits_all <- 
  BB_Species |> 
  merge(SeedNumber, by = "species") |>
  merge(SeedGermination, by = "species") |>
  merge(PDensity, by = "species") |>
  merge(Wetmass, by = "species") |>
  merge(Dispersal, by = "species") |>
  merge(Lambda, by = "species")
  
unique(traits_all$species)  #check the number of species with data for all traits

print(traits_all)

```


# 2.2 - Final trait table

Calculate the mean and standard deviation for each species for each trait.

```{r}

#select only columns with the values for final table
simplified_traits <- traits_all[,c("species","seed_number","germination_perc","density_sqMeter","wetmass_g","mortality","mean_dist_m","max_dist_m","lambdaVals")]


#create final table
final <- simplified_traits[seq_along(unique(simplified_traits$species)),]

#delete data from table
final[TRUE, TRUE] = NA

#add species to column
final$species = unique(simplified_traits$species)

#duplicate trait columns
final <- cbind(final, final[, c(2:9)])

#change names
colnames(final)[2:9] = paste0(colnames(final)[2:9], "_mean")
colnames(final)[10:17] = paste0(colnames(final)[10:17], "_std")


#add traits

for (row in seq_len(nrow(final))) { #for each row (aka species) in the final table
  sp = final$species[row] #get the species name
  sp_rows = simplified_traits[simplified_traits$species == sp,] #get the rows for that species from the simplified_traits dataframe

  for (col in colnames(sp_rows)) { #for each column (aka trait)
    if (col == "species") next #skip species column

    if (length(unique(sp_rows[,col])) == 1) { #if there is only one unique value for that trait
      final[row, paste0(col, "_mean")] = sp_rows[1,col] #the mean is that number
final[row, paste0(col, "_std")] = 0 #the std is 0
    } else { #if there are multiple values
      if (is.character(sp_rows[,col])) { #must be numeric, else stop function
        stop("You have a character value in this column!")
      } else { #if is numeric
        final[row, paste0(col, "_mean")] = mean(sp_rows[,col]) #calculate mean
final[row, paste0(col, "_std")] = sd(sp_rows[,col]) #calculate std
      }
    }
  }
}

print(final)

#write.csv(final, "Results/trait_data_vfinal.csv")

```


# 3 - Occurrence data


```{r}
sp_list <- data.frame(unique(traits_all$species)) #get species names included in the complete data
names(sp_list) <- 'species' #change column name

```


## Calculate occurrences in Bavaria for each species

Used the same file (GBIF_bav) that was downloaded in Chapter 1.

```{r}

gbif_bav <- merge(sp_list, GBIF_bav, by = "species") #get occurrence data from Bavaria for the species included in the complete data

library(stats) #for aggregate function

occur_sp_bav <- aggregate(numberOfOccurrences ~ species, data=gbif_bav, sum) #calculate number of occurrences for each species

names(occur_sp_bav)[names(occur_sp_bav) == 'numberOfOccurrences'] <- 'numberOfOccurrences_bav' #change column name

```

## Calculate occurrences in BW for each species

Used the same file (GBIF_bav) that was downloaded in Chapter 1.

```{r}

gbif_bw <- merge(sp_list, GBIF_baden, by = "species") #get occurrence data from BW for the species included in the complete data

occur_sp_bw <- aggregate(numberOfOccurrences ~ species, data=gbif_bw, sum) #calculate number of occurrences for each species

names(occur_sp_bw)[names(occur_sp_bw) == 'numberOfOccurrences'] <- 'numberOfOccurrences_bw' #change column name

```

## Calculate occurrences in Bavaria + BW for each species

```{r}

occur <- merge(occur_sp_bav, occur_sp_bw, by = "species", all=TRUE) #merge occurrences from Bavaria and BW

occur$numberOfOccurrences <- rowSums(cbind(occur$numberOfOccurrences_bav,occur$numberOfOccurrences_bw),na.rm=TRUE) #sum occurrences

print(occur)

write.csv(occur, "Results/Occurrences_bav&bw.csv") #download occurrence file

```


## Calculate occurrences worldwide for each species

Go to https://www.gbif.org/occurrence/search?occurrence_status=present&q= to download the species list with occurrences, and use the following filters: 
  - "Scientific name" - "Tracheophyta"
  - "Location" - "Including coordinates"
  
Download - option "species list"

```{r}

GBIF_world <- read.csv(unzip("Data/SpeciesList_Tracheophyta_worldwide.zip"), sep = "\t", quote = "") #read file

gbif_world <- merge(sp_list, GBIF_world, by = "species") #get occurrence data worldwide for the species included in the complete data

occur_sp_ww <- aggregate(numberOfOccurrences ~ species, data=gbif_world, sum) #calculate number of occurrences for each species

print(occur_sp_ww)

write.csv(occur_sp_ww, "Results/Occurrences_worldwide.csv") #download occurrence file

```

<br><br>
<div class="tocify-extend-page" data-unique="tocify-extend-page" style="height: 0;"></div>
